---
title: "Estimating temporary populations: a systematic review of the empirical literature"
subtitle: "Supplementary Information"
author: "Radoslaw Panczak, Elin Charles-Edwards, Jonathan Corcoran"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::html_document2:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    # toc_float: yes
link-citations: yes
csl: bibliography/springerplus.csl
bibliography:
- bibliography/Temporary Populations.bib
header-includes: 1
  - \pagenumbering{gobble}
nocite: | 
  @censusofenglandandwales1921Workplaces1925, 
  @thornthwaiteLouisvilleKentuckyStudy1929,
  @breeseDaytimePopulationCentral1947,
  @newyorkregionalplanassociationPersonsVehiclesEntering1949,
  @seattlecityplanningcommissionDaytimeNighttimePopulation1951,
  @foleyDailyMovementPopulation1952,
  @instituteforresearchinsocialscienceuniversityofnorthcarolinaPopulationDistributionSpatial1952,
  @menzlerEstimateDaytimePopulation1952,
  @sharpNonResidentialPopulationCentral1955,
  @bureauofthecensusPopulationEstimatesSurvival1956,
  @weirSurveyDaytimePopulation1960,
  @ericksonMeasuringResortPopulation1961,
  @wurtelePopulationDynamicsFinal1968,
  @goldschmidtDemoflushEstimatingPopulation1976,
  @fultonEstimatingDaytimePopulation1984,
  @goberGeographyNonpermanentResidence1984,
  @kavanaughMethodEstimatingDaytime1990,
  @banzPredictionTimedependentPopulation1992,
  @nelsonEstimatingFunctionalPopulation1992,
  @stutzChartingUrbanSpacetime1992,
  @akkermanUrbanHouseholdPattern1995,
  @galvezMeasurementFloridaTemporary1996,
  @bellPatternsTemporaryMobility1998,
  @roddisMetropolitanActivityProfiles1996,
  @taylorMeasuringShortTerm1998,
  @yongliResearchStatisticalModels1998,
  @warchivkerTurmoilAboriginalEnumeration2000,
  @lambEstimatingSeasonalPopulation1999,
  @happelCountingSnowbirdsImportance2002,
  @mcphersonDayNightPopulation2003,
  @smithNationalPopulationData2005,
  @sleeterEstimatingDaytimeNighttime2006,
  @aholaSpatiotemporalPopulationModel2007,
  @bhaduriLandScanUSAHighresolution2007,
  @collinsDaytimePopulationTracking2007,
  @needhamUsingGISModel2007,
  @taubenbockLinkingStructuralUrban2007,
  @monmouthcountyplanningboardSummerCoastalPopulation2008,
  @lauGISBasedStochastic2009,
  @campbellDevelopingExtensibleSolution2010,
  @freireModelingSpatiotemporalDistribution2010,
  @horanontEstimateAmbientPopulation2010,
  @rigall-i-torrentEstimatingOvernightFacto2010,
  @bellComparingMethodsEstimation2011,
  @bengtssonImprovedResponseDisasters2011,
  @charles-edwardsModellingFluxEstimation2011,
  @kobayashiVisualizingDiurnalPopulation2011,
  @sekimotoPFlowReconstructingPeople2011,
  @swansonEstimatingFactoPopulation2011,
  @walkerAgentbasedPopulationModel2012,
  @charles-edwardsEstimatingServicePopulation2013,
  @hugoTimeTideMoving2013,
  @markhamNewDirectionsIndigenous2013,
  @mckenzieCommuteradjustedPopulationEstimates2013,
  @officefornationalstatistics2011CensusWorkday2013,
  @teradaPopulationEstimationTechnology2013,
  @araImpactTemporalPopulation2014,
  @devilleDynamicPopulationMapping2014,
  @fehrpeersNapaCountyTravel2014,
  @gaoAssessingSocialEconomic2014,
  @graebertNorthwestMichiganSeasonal2014,
  @himotoModelSpatiotemporalDistribution2014,
  @gregerSpatioTemporalBuildingPopulation2015,
  @hodurServicePopulationEstimates2015,
  @martinDevelopingFlexibleFramework2015,
  @qiModelingSpatialDistribution2015,
  @khakpourAttractionbasedCellularAutomaton2016,
  @khodabandelouPopulationEstimationMobile2016,
  @lwinSpaceTimeMultiple2016,
  @mckenzieDemographicsFireRisk2016,
  @wilsonRapidRealTimeAssessments2016,
  @xuContextawareRealtimePopulation2016,
  @adamiakSeasonalResidenceCounterurbanization2017,
  @batistaesilvaSpatiotemporalMappingPopulation2017,
  @jarvEnhancingSpatialAccuracy2017,
  @kashiyamaOpenPFLOWCreation2017,
  @kimModellingDaytimePopulation2017,
  @kontokostaUrbanPhenologyRealtime2017,
  @maModelingHourlyDistribution2017,
  @mortonSimulationApproachModeling2017,
  @thomasUseMobileDevice2017,
  @stathakisSeasonalPopulationEstimates2018,
  @batranInferencingHumanSpatiotemporal2018,
  @boeingEstimatingLocalDaytime2018,
  @chenFinegrainedPredictionUrban2018,
  @esriMethodologyStatement20182018,
  @liuMappingHourlyDynamics2018,
  @picornellPopulationDynamicsBased2018,
  @thakurMethodologyAccountSeasonal2018,
  @zagattiTripWorkEstimation2018,
  @zinEstimationOriginatingDestinationTrips2018,
  @crolsQuantifyingAmbientPopulation2019,
  @edmondsonMakingItCount2019,
  @harrisWashingtonCountyTemporary2019,
  @xuUnravelingEnvironmentalJustice2019,
  @yunImplementationFloatingPopulation2019
---

# Supplementary Text S1: Search strategy

We searched four major databases of scientific literature: arXiv, Google Scholar, Scopus and Web of Science. 

## arXiv 

arXiv was searched on the 21 December 2017 using web interface. We used all search terms defined in the main text. Since there was no possibility to set up automatic updates we monitored the source manually until the day of submission of the manuscript. 

## Google Scholar 

Google Scholar was searched between 8 and 9 November 2017 using Publish or Perish software (Harzing 2007, available from https://harzing.com/resources/publish-or-perish). Due to very large amount of retried records we excluded search terms "working population",	"mobile population", "service population",	"floating population" and "seasonal population". We updated the search for the "spatiotemporal population" keyword on 29 November 2017. Because of the limitations of the data retrieval we searched for each of the keywords separately and limited the number of years of publication to achieve full coverage. The results were then monitored by scrutinizing email update service set up for each search. 

## Scopus

Scopus was searched on the 7 November 2017 using web interface. We searched for primary results and "secondary documents". We excluded subject areas not related to the topic ("Medicine", "Biochemistry, Genetics and Molecular Biology", "Chemistry", "Psychology", "Immunology and Microbiology"). We updated the search for the "spatiotemporal population" keyword on 24 November 2017. The results were then monitored by scrutinizing RSS service feed with updates for each search. 

## Web Of Science 

Web Of Science was searched on the 7 November 2017 using web interface. We searched for primary results and "secondary documents". We excluded categories not related to the topic ("TROPICAL MEDICINE", "NEUROSCIENCES", "PARASITOLOGY", "MATHEMATICS", "ENVIRONMENTAL SCIENCES", "DEVELOPMENTAL BIOLOGY", "MEDICINE RESEARCH EXPERIMENTAL", "MARINE FRESHWATER BIOLOGY", "BIOCHEMICAL RESEARCH METHODS", "HEALTH CARE SCIENCES SERVICES", "MULTIDISCIPLINARY SCIENCES", "SOIL SCIENCE", "BIODIVERSITY CONSERVATION", "WATER RESOURCES", "VIROLOGY", "ZOOLOGY", "OPHTHALMOLOGY", "EVOLUTIONARY BIOLOGY", "FORESTRY", "IMMUNOLOGY", "OCEANOGRAPHY", "COMPUTER SCIENCE THEORY METHODS", "ENGINEERING CIVIL", "BIOLOGY", "PUBLIC ENVIRONMENTAL OCCUPATIONAL HEALTH", "VETERINARY SCIENCES", "MATHEMATICAL COMPUTATIONAL BIOLOGY", "ANATOMY MORPHOLOGY", "FISHERIES", "OPTICS", "PLANT SCIENCES", "BIOTECHNOLOGY APPLIED MICROBIOLOGY", "GEOSCIENCES MULTIDISCIPLINARY", "ORNITHOLOGY", "ONCOLOGY", "BIOCHEMISTRY MOLECULAR BIOLOGY", "IMAGING SCIENCE PHOTOGRAPHIC TECHNOLOGY", "GENETICS HEREDITY", "PHYSICS FLUIDS PLASMAS", "ENTOMOLOGY", "BEHAVIORAL SCIENCES", "CARDIAC CARDIOVASCULAR SYSTEMS", "ORTHOPEDICS", "COMPUTER SCIENCE CYBERNETICS", "SPORT SCIENCES", "COMPUTER SCIENCE HARDWARE ARCHITECTURE", "AGRONOMY", "GEOCHEMISTRY GEOPHYSICS", "MEDICAL INFORMATICS", "INFECTIOUS DISEASES", "BIOPHYSICS", "PSYCHOLOGY EXPERIMENTAL", "PHYSIOLOGY", "LIMNOLOGY", "MECHANICS", "CHEMISTRY PHYSICAL", "ENGINEERING BIOMEDICAL", "NEUROIMAGING", "NANOSCIENCE NANOTECHNOLOGY", "ENGINEERING ELECTRICAL ELECTRONIC", "TOXICOLOGY", "PHYSICS CONDENSED MATTER", "COMPUTER SCIENCE INTERDISCIPLINARY APPLICATIONS", "CLINICAL NEUROLOGY", "CHEMISTRY MULTIDISCIPLINARY", "METEOROLOGY ATMOSPHERIC SCIENCES", "PHYSICS ATOMIC MOLECULAR CHEMICAL", "REHABILITATION", "PSYCHOLOGY", "RADIOLOGY NUCLEAR MEDICINE MEDICAL IMAGING", "MEDICINE GENERAL INTERNAL", "ACOUSTICS", "MICROBIOLOGY", "AGRICULTURE MULTIDISCIPLINARY", "ENGINEERING MECHANICAL", "CELL BIOLOGY"). We updated the search for the "spatiotemporal population" keyword on 24 November 2017. The results were then monitored by scrutinizing RSS service feed with updates for each search. 

## Reviews of the literature 

Blondel, V. D., Decuyper, A., & Krings, G. (2015). A survey of results on mobile phone datasets analysis. EPJ Data Science, 4(1), 10. https://doi.org/10.1140/epjds/s13688-015-0046-0

Daas, P. J. H., Puts, M., Tennekes, M., & Priem, A. (2014). Big Data as a Data Source for Official Statistics: experiences at Statistics Netherlands. In Proceedings of Statistics Canada Symposium 2014 Beyond traditional survey taking: adapting to a changing world. Gatineau, Canada: Statistics Canada.

Hughes, C., Zagheni, E., Abel, G. J., Sorichetta, A., Wisniowski, A., Weber, I., & Tatem, A. J. (2016). Inferring migrations, traditional methods and new approaches based on mobile phone, social media, and other big data: feasibility study on inferring (labour) mobility and migration in the European Union from big data and social media data. (p. 41). Directorate-General for Employment, Social Affairs and Inclusion (European Commission). Retrieved from https://publications.europa.eu/en/publication-detail/-/publication/1f66f928-f307-4c1f-9bec-fde0d2008c69/language-en

Li, J., Xu, L., Tang, L., Wang, S., & Li, L. (2018). Big data in tourism research: A literature review. Tourism Management, 68, 301–323. https://doi.org/10.1016/j.tourman.2018.03.009

Liao, C., Brown, D., Fei, D., Long, X., Chen, D., & Che, S. (2018). Big data-enabled social sensing in spatial analysis: Potentials and pitfalls. Transactions in GIS. https://doi.org/10.1111/tgis.12483

Naboulsi, D., Fiore, M., Ribot, S., & Stanica, R. (2016). Large-Scale Mobile Traffic Analysis: A Survey. IEEE Communications Surveys Tutorials, 18(1), 124–161. https://doi.org/10.1109/COMST.2015.2491361

Martí, P., Serrano-Estrada, L., & Nolasco-Cirugeda, A. (2018). Social Media data: Challenges, opportunities and limitations in urban studies. Computers, Environment and Urban Systems. https://doi.org/10.1016/j.compenvurbsys.2018.11.001

Williams, S. (2016). Statistical uses for mobile phone data: literature review (ONS methodology working paper series No. 8). Newport: Office for National Statistics. Retrieved from https://www.ons.gov.uk/methodology/methodologicalpublications/generalmethodology/onsworkingpaperseries/onsmethodologyworkingpaperseriesno8statisticalusesformobilephonedataliteraturereview





# Supplementary Text S2: Full list of studies included in the review.

<div id="refs"></div>

# Supplementary Table S2: Main characteristics of the studies.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(pacman)
p_load(tidyverse, readxl, kableExtra)

```

```{r data-main, include=FALSE}
main <- read_excel("./data/TEMPO_SR_2019-10-22.xlsx", 
                   sheet = "Main") %>% 
  replace(is.na(.), "")
```

## Source, type, purpose, main data and methods 

```{r}
main %>% 
  select(FA_surname, Year_pub, 
         Source, Pub_type, 
         Purpose_1, Purpose_2, 
         Type_1, Type_2, Type_3, Type_4, 
         Method) %>% 
  kable(col.names = c("Author",
                      "Year of publication",
                      "Source",
                      "Type of publication",
                      "Purpose (1)",  
                      "Purpose (2)",  
                      "Main data (1)",
                      "Main data (2)",
                      "Main data (3)",
                      "Main data (4)",
                      "Methods")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

## Coverage, country, region, size of data & population, validation, software, availability of estimates

```{r}
main %>% 
  select(FA_surname, Year_pub,
         National, Country_1, Country_2, 
         Region_type, Region,
         Data_size_users, Est_pop_size, 
         Validation, Software_avail, Est_avail) %>% 
  kable(col.names = c("Author",
                      "Year of publication",
                      "National coverage", 
                      "Country (1)", "Country (2)", 
                      "Type of Region", 
                      "Name of Region",
                      "Size of population (data)", 
                      "Size of population (estimated)", 
                      "Was validation of the estimates carried out?", 
                      "Is software used for the estimation avaiable?", 
                      "Are the estimates avaiable for download?")) %>%  
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

## Temporal & spatial units 

```{r}
main %>% 
  select(FA_surname, Year_pub,
         Day, Est_tunit, Est_tunit_num, 
         Est_unit_type_1, Est_unit_type_2	, Est_spat_num_1, Est_spat_num_2) %>% 
  kable(col.names = c("Author",
                      "Year of publication",
                      "Daytime estimates", 
                      "Type of temporal unit", 
                      "Number of temporal unit", 
                      "Spatial unit for estimated populations (1)",
                      "Spatial unit for estimated populations (2)",
                      "Number of spatial units for estimated population (1)", 
                      "Number of spatial units for estimated population (2)")) %>%  
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

# Supplementary Table S3: Applications & related studies.

```{r data-sup, include=FALSE}
applications <- read_excel("./data/TEMPO_SR_2019-10-22.xlsx", sheet = "Related") %>% 
  replace(is.na(.), "")
```

```{r}
applications %>% 
  kable(col.names = c("Author",
                      "Year of publication",
                      "Applications and related studies")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
